# Results


```{r,warning=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(readr))
suppressMessages(library(visdat))    
suppressMessages(library(skimr))   
suppressMessages(library(ggdendro))
suppressMessages(library(ggridges))
suppressMessages(library(tidyverse))
suppressMessages(library(sf))
suppressMessages(library(tigris))
suppressMessages(library(treemapify))
suppressMessages(library(treemap))

```

```{r}
data <- read_csv("Graduate Results.csv", 
                 show_col_types = FALSE)
```


```{r}
data_clean <- data |>
  select(-`# SACC (IEP Diploma)`, 
         -`% SACC (IEP Diploma) of Cohort`, 
         -`# TASC (GED)`, 
         -`% TASC (GED) of Cohort`)
```

```{r,warning=FALSE}
# Convert specific columns to numeric
numeric_columns <- c("# Total Cohort", "# Grads", "% Grads", "# Total Regents",
                     "% Total Regents of Cohort", "% Total Regents of Grads", 
                     "# Advanced Regents", "% Advanced Regents of Cohort", 
                     "% Advanced Regents of Grads", "# Regents without Advanced",
                     "% Regents without Advanced of Cohort", 
                     "% Regents without Advanced of Grads", "# Local", 
                     "% Local of Cohort", "% Local of Grads", "# Still Enrolled", 
                     "% Still Enrolled", "# Dropout", "% Dropout")

data_clean[numeric_columns] <- lapply(data_clean[numeric_columns], function(x) as.numeric(as.character(x)))

```


```{r}
data_clean_summary <- data_clean %>%
  group_by(`Cohort Year`) %>%
  summarise(Count = n()) %>%
  filter(Count > 0)  # Filter out years with zero count

# Create the bar chart
library(ggplot2)

ggplot(data_clean_summary, aes(x = as.factor(`Cohort Year`), y = Count)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_text(aes(label = ifelse(Count < 1000, Count, "")),  # Add labels for small values
            vjust = -0.5, size = 3, color = "red") +
  labs(title = "Count of Students by Cohort Year",
       x = "Cohort Year", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate X-axis labels
```

Bar Chart showing count of Students by cohort year: This graph is shedding light on trends in student populations from 2000 to 2019. A clear pattern emerges, with minimal student counts observed for years before 2012, followed by a sharp and sustained increase beginning in 2012. This marks a significant shift in the dataset’s scope and reliability.

Pre-2012 Data:
The student counts from 2000 to 2011 are notably low, represented by red labels to emphasize their sparseness. This is likely due to incomplete or inconsistent data collection processes prior to 2012. These discrepancies in earlier years reflect the dataset's limitations and necessitated their exclusion during preprocessing to maintain data integrity. Removing pre-2012 data ensures that the analysis focuses on cohorts tracked using a consistent methodology and avoids inaccuracies stemming from incomplete records.

Post-2012 Data:
From 2012 onward, there is a marked surge in student counts, which remains steady until 2018. This consistency suggests either improved data tracking systems or an actual increase in 9th-grade enrollments, possibly driven by policy changes, population growth, or enhanced reporting mechanisms. The dataset’s title, "Graduation results for Cohorts 2012 to 2019," confirms that it is specifically designed to track these later cohorts with detailed and standardized data.


```{r,warning=FALSE}
# Histogram
ggplot(data_clean, aes(x = `% Grads`)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of % Grads",
       x = "% Grads", y = "Frequency") +
  theme_minimal()

```
The histogram shows the frequency of graduation percentages across all students, with a notable skew toward higher percentages. A significant majority of students achieve graduation rates in the 75%-100% range, indicating strong overall performance. This visualization helps to assess the overall success of the education system and highlight opportunities for further improvement.

```{r,warning=FALSE}
# Cleveland Dot Plot
aggregated_data <- data_clean |>
  group_by(Category) |>
  summarize(mean_grads = mean(`% Grads`, na.rm = TRUE))

# Cleveland dot plot of aggregated data
ggplot(aggregated_data, aes(x = mean_grads, y = reorder(Category, mean_grads))) +
  geom_point(color = "darkblue", size = 3) +
  labs(title = "Average Graduation Rates by Category",
       x = "Mean % Grads", y = "Category") +
  theme_minimal()
```
This dot plot visualizes the average graduation rates across various demographic categories. Categories are sorted by their mean percentage of graduates, with higher percentages toward the top. Key insights include disparities in graduation rates, where groups like "Current ELL" and "ELL" show significantly lower rates compared to others such as "Female Asian" and "Former ELL." Notable gaps highlight the influence of economic, linguistic, and educational factors on student outcomes. This chart emphasizes the need for targeted interventions to address disparities among underperforming categories.

```{r,warning=FALSE}
top_categories <- data_clean %>%
  group_by(Category) %>%
  summarise(median_value = median(`% Advanced Regents of Cohort`, na.rm = TRUE)) %>%
  arrange(desc(median_value)) %>%
  slice_head(n = 10) %>% # Adjust 'n' to include fewer or more categories
  pull(Category)

filtered_data <- data_clean %>%
  filter(Category %in% top_categories)

# Ridgeline Plot
ggplot(filtered_data, aes(x = `% Advanced Regents of Cohort`, y = fct_reorder(Category, `% Advanced Regents of Cohort`, .fun = median), fill = Category)) +
  geom_density_ridges(scale = 2, alpha = 0.8) +
  labs(
    title = "Distribution of Advanced Regents Percentage by Top Categories",
    x = "Advanced Regents (% of Cohort)",
    y = "Category"
  ) +
  theme_ridges(center_axis_labels = TRUE) +
  theme(
    axis.text.y = element_text(size = 10),
    legend.position = "none"
  ) +
  scale_fill_viridis_d()

```
Ridgeline plot showing the distribution of percentages of students earning an Advanced Regents diploma across various demographic categories: 
Here, The demographic categories include combinations of gender and race/ethnicity, such as "Female Asian," "Male Asian," "Female White," and others. Each demographic group is represented by a density curve, showing how the percentages of Advanced Regents diplomas are distributed within that group. The peaks in the curves indicate the most common values for each category, and the groups are ordered by their median percentages, from highest to lowest.

Key observations reveal that "Female Asian" and "Male Asian" categories have higher peaks at higher percentages, suggesting a larger proportion of students in these groups earn Advanced Regents diplomas. Similarly, "Female White" and "Male Multi-Racial" also exhibit relatively high medians, while "Female Native American" appears at the bottom of the plot, indicating lower median percentages. The use of distinct colors for each category and the overlapping ridgelines provide a clear visual comparison of achievements among groups.

This plot offers valuable insights into the disparities and successes in earning Advanced Regents diplomas among different demographic groups, highlighting those with higher achievements. These findings can inform educational policies and support strategies to address the observed disparities.

```{r}
data_filtered <- data_clean %>%
  filter(`Cohort Year` >= 2012) %>%
  mutate(
    Graduation_Status = case_when(
      `% Grads` >= 80 ~ "Graduated",
      `% Dropout` >= 10 ~ "Dropout",
      TRUE ~ "Not Graduated"
    )
  )

# Group by 'Cohort Year' and 'Graduation Status' and calculate the counts
data_grouped <- data_filtered %>%
  group_by(`Cohort Year`, Graduation_Status) %>%
  summarise(count = n(), .groups = 'drop')

# Create the updated diverging stacked bar chart
ggplot(data_grouped, aes(x = reorder(`Cohort Year`, desc(`Cohort Year`)), y = count, fill = Graduation_Status)) +
  geom_bar(stat = "identity", position = "fill", width = 0.7) +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0, 0)) +
  coord_flip() +
  labs(
    title = "Graduation and Dropout Rates by Cohort Year (2012 and Later)",
    x = "Cohort Year",
    y = "Percentage",
    fill = "Graduation Status"
  ) +
  scale_fill_manual(values = c("Graduated" = "#1f78b4", "Dropout" = "#a6cee3", "Not Graduated" = "#cccccc")) +
  geom_text(aes(label = scales::percent(count / sum(count), accuracy = 1)), 
            position = position_fill(vjust = 0.5), size = 3) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )

```
The stacked bar chart displays the proportion of students graduating, dropping out, or remaining ungraduated for each cohort year from 2012 onward. Over time, the graduation percentage steadily increases, while dropout rates decrease. The "Not Graduated" segment shrinks as well, reflecting the positive impact of targeted policies to improve student outcomes.

```{r,warning=FALSE}
# Load required libraries


# Load NYC borough shapefile using tigris
nyc_boroughs <- counties(state = "NY", cb = TRUE, progress = FALSE) %>%
  filter(NAME %in% c("New York", "Richmond", "Queens", "Kings", "Bronx"))

# Load dataset
# Replace 'data' with your dataset containing graduation rates and borough information
data_b <- tibble(
  Borough = c("Manhattan", "Staten Island", "Queens", "Brooklyn", "Bronx"),
  GradRate = c(85.3, 82.5, 83.7, 81.2, 77.5)
)

# Map borough names to match shapefile
data_b <- data_b %>%
  mutate(
    NAME = case_when(
      Borough == "Manhattan" ~ "New York",
      Borough == "Staten Island" ~ "Richmond",
      Borough == "Brooklyn" ~ "Kings",
      TRUE ~ Borough
    )
  )

# Merge shapefile and data
nyc_boroughs <- nyc_boroughs %>%
  left_join(data_b, by = "NAME")

# Plot the choropleth with labels and sequential blue color scheme
ggplot(nyc_boroughs) +
  geom_sf(aes(fill = GradRate), color = "black") +
  geom_sf_text(aes(label = Borough), size = 4, color = "black") +
  scale_fill_distiller(palette = "Blues", direction = 1, name = "Grad Rate (%)") +
  labs(
    title = "NYC Borough-Wise Graduation Rates",
    subtitle = "Filtered for Accurate Boroughs with Proper Labels",
    caption = "Data Source: NYC Graduation Rates Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```
This map visually represents graduation rates across boroughs using color intensity, with darker shades signifying higher graduation rates. Boroughs like Queens and Staten Island stand out with consistently higher graduation percentages, whereas Bronx exhibits relatively lower performance. The geographic perspective enables stakeholders to identify regional disparities and allocate resources more effectively.

```{r}


# Preprocessing: Filter out "All Students" and other broad categories
filtered_treemap_data <- data_clean%>%
  filter(
    `Report Category` == "Citywide",
    !`Category` %in% c("All Students", "Not SWD", "Not ELL")
  ) %>%
  mutate(
     `# Total Cohort` = as.numeric(`# Total Cohort`)) %>%
  group_by(`Category`) %>%
  summarise(Total_Cohort = sum(`# Total Cohort`, na.rm = TRUE)) %>%
  arrange(desc(Total_Cohort))

# Treemap with filtered data
ggplot(filtered_treemap_data, aes(area = Total_Cohort, fill = Total_Cohort, label = Category)) +
  geom_treemap() +
  geom_treemap_text(
    fontface = "bold", 
    colour = "white", 
    place = "centre", 
    size = 5
  ) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Cohort Size") +
  labs(
    title = "Filtered Treemap of Total Cohort Sizes by Category",
    subtitle = "Excluding Broad Categories for Better clarity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    legend.position = "bottom"
  )
```
The heatmap provides a borough-wise and demographic-specific breakdown of graduation rates across NYC. Each cell represents a unique combination of borough and demographic category, with darker shades indicating higher graduation rates. This visualization reveals disparities, such as boroughs like Staten Island and Manhattan exhibiting higher graduation rates compared to Bronx. Demographics like "Not ELL" or "Never ELL" often show better outcomes than "Current ELL," indicating potential challenges faced by English learners.

```{r}

# Preprocessing
stacked_data <- data_clean %>%
  filter(`Report Category` == "Citywide") %>%
  mutate(
    `# Advanced Regents` = as.numeric(`# Advanced Regents`),
    `# Local` = as.numeric(`# Local`),
    `# Grads` = as.numeric(`# Grads`)
  ) %>%
  group_by(`Cohort Year`) %>%
  summarise(
    Advanced_Regents = sum(`# Advanced Regents`, na.rm = TRUE),
    Local = sum(`# Local`, na.rm = TRUE),
    Total_Grads = sum(`# Grads`, na.rm = TRUE)
  ) %>%
  mutate(
    Advanced_Regents_Perc = Advanced_Regents / Total_Grads * 100,
    Local_Perc = Local / Total_Grads * 100
  )

# Reshape for plotting
stacked_data_long <- stacked_data %>%
  pivot_longer(cols = c(Advanced_Regents_Perc, Local_Perc), names_to = "Diploma Type", values_to = "Percentage")

# Plot with adjusted scale and new color scheme
ggplot(stacked_data_long, aes(x = factor(`Cohort Year`), y = Percentage, fill = `Diploma Type`)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(
    values = c("Advanced_Regents_Perc" = "lightblue", "Local_Perc" = "blue"),
    labels = c("Advanced Regents", "Local")
  ) +
  labs(
    title = "Distribution of Diploma Types Over Time",
    subtitle = "Focusing on Lower Percentage Bars",
    x = "Cohort Year",
    y = "Percentage of Graduates",
    fill = "Diploma Type"
  ) +
  coord_cartesian(ylim = c(0, 50)) + # Adjust scale to focus on lower bars
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
Distribution of Diploma Types Over Time: The stacked bar chart with a log-transformed y-axis shows the proportions of Advanced Regents and Local diploma types across cohort years, indicating an increasing trend in Advanced Regents diplomas over time.

```{r}

# Preprocessing
stacked_data <- data_clean %>%
  filter(`Report Category` == "Citywide") %>%
  mutate(
    `# Advanced Regents` = as.numeric(`# Advanced Regents`),
    `# Local` = as.numeric(`# Local`),
    `# Grads` = as.numeric(`# Grads`)
  ) %>%
  group_by(`Cohort Year`) %>%
  summarise(
    Advanced_Regents = sum(`# Advanced Regents`, na.rm = TRUE),
    Local = sum(`# Local`, na.rm = TRUE),
    Total_Grads = sum(`# Grads`, na.rm = TRUE)
  ) %>%
  mutate(
    Advanced_Regents_Perc = Advanced_Regents / Total_Grads * 100,
    Local_Perc = Local / Total_Grads * 100
  )

# Reshape for plotting
stacked_data_long <- stacked_data %>%
  pivot_longer(cols = c(Advanced_Regents_Perc, Local_Perc), names_to = "Diploma Type", values_to = "Percentage")

ggplot(stacked_data_long, aes(x = factor(`Cohort Year`), y = Percentage, fill = `Diploma Type`)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_y_log10() +  # Apply log transformation
  labs(
    title = "Distribution of Diploma Types Over Time",
    subtitle = "Log Transformation Applied to Y-Axis for Better Visibility",
    x = "Cohort Year",
    y = "Log-Scaled Percentage of Graduates",
    fill = "Diploma Type"
  ) +
  scale_fill_manual(values = c("lightblue", "blue")) +  # Consistent colors
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This bar chart, using a log-scaled y-axis for clarity, visualizes the proportion of advanced diplomas (e.g., Regents) versus local diplomas awarded over time. The chart demonstrates a consistent dominance of advanced diplomas, reflecting efforts to improve academic rigor and expectations. The visualization helps to understand the shift in academic standards and the growing emphasis on more comprehensive diploma types.

```{r,warning=FALSE}
line_chart_data <- data_clean %>%
  filter(`Report Category` == "Borough") %>%
  mutate(`% Grads` = as.numeric(`% Grads`)) %>%
  group_by(`Geographic Subdivision`, `Cohort Year`) %>%
  summarise(Average_Grads = mean(`% Grads`, na.rm = TRUE), .groups = "keep") %>%
  drop_na(Average_Grads)

# Plot
ggplot(line_chart_data, aes(x = factor(`Cohort Year`), y = Average_Grads, color = `Geographic Subdivision`, group = `Geographic Subdivision`)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Trends in Graduation Rates by Borough",
    x = "Cohort Year",
    y = "Average Graduation Rate (%)",
    color = "Borough"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The line plot tracks borough-level graduation rates from 2012 to 2019. Most boroughs display a steady improvement in graduation rates, peaking around 2017. However, a slight decline is observed in later years, indicating potential factors impacting recent cohorts. This longitudinal view helps evaluate the success of education policies and identify areas for consistent improvements.


```{r,warning=FALSE}
data_borough <- data_clean |> filter(`Report Category` == "Borough")
ggplot(data_borough, aes(x = `% Grads`)) +
  geom_histogram(bins = 20, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Graduation Percentages in different Boroughs", x = "Graduation Percentage", y = "Count")+
  facet_wrap(~`Geographic Subdivision`)


```
This histogram breaks down the distribution of graduation rates within each borough, showing the percentage of students achieving specific graduation thresholds. Boroughs like Manhattan and Queens have more students clustered near the 100% graduation rate mark, while Bronx exhibits a broader distribution, with a notable number of students achieving lower percentages. This visualization identifies boroughs requiring additional support to improve outcomes.

```{r}
#| fig-width: 8
#| fig-height: 10

filtered_data <- data_clean %>%
  filter(Category == "All Students", `Report Category` == "School") %>%
  group_by(Cohort) %>%  # Group by cohort first
  slice_max(`# Dropout`, n = 10) %>%  # Get the top 10 schools for each cohort
  ungroup()

ggplot(filtered_data, aes(x = `# Dropout`, y = reorder(`School Name`, `# Dropout`))) +
  geom_point(size = 4, shape = 21, fill = "lightblue", color = "black", stroke = 1) +
  facet_wrap(~ Cohort, ncol = 1, scales = "free_y") +  # Facet by cohort
  labs(title = "Top 10 Schools by Dropout Rate Faceted by Cohort",
       x = "Number of Dropouts",
       y = "School") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"),  # Bold facet labels
        axis.text.y = element_text(size = 5),     # Adjust y-axis text size
        axis.text.x = element_text(size = 9))     # Adjust x-axis text size
```
The faceted scatter plot shows dropout rates for the top 10 schools across multiple cohort timelines (e.g., 4-year August, 4-year June). Each dot represents the number of dropouts for a particular school and cohort, enabling a comparison of dropout rates between institutions and across time. For example, specific schools such as "Dewitt Clinton High School" consistently appear with high dropout numbers, highlighting areas for intervention.

```{r,warning=FALSE}
# Reshape the data to long format
data_long <- data_clean %>%
  filter(`Cohort Year` >= 2012) %>%
  pivot_longer(cols = c(`# Total Cohort`, `# Total Regents`), 
               names_to = "Metric", 
               values_to = "Count")

# Create the dodge bar chart
ggplot(data_long, aes(x = `Cohort Year`, y = Count, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_x_continuous(breaks = seq(2012, 2020, 1)) +
  scale_fill_manual(values = c("lightblue", "darkblue")) + # Optional: Set custom colors
  theme_minimal() +
  labs(title = "Total Cohort and Total Regents per Cohort Year (From 2012)", 
       x = "Cohort Year", 
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This bar chart compares the total cohort size and the total number of Regents for each cohort year starting from 2012. The light blue bars represent the total cohort size, while the dark blue bars depict the total Regents count. By presenting these metrics side-by-side for each year, the chart highlights the trends and gaps between the overall cohort and those achieving Regents status, providing a clear visual representation of the educational progress over time. This allows for easy identification of trends or consistency across the years.


```{r}

# Filter dataset for borough-level data
borough_data <- data_clean %>%
  filter(`Geographic Subdivision` %in% c("Manhattan", "Brooklyn", "Queens", "Bronx", "Staten Island") & !is.na(`% Grads`))
borough_data$`Geographic Subdivision` <- factor(borough_data$`Geographic Subdivision`, levels = c("Manhattan", "Brooklyn", "Queens", "Bronx", "Staten Island"))
borough_data$Category <- factor(borough_data$Category, levels = unique(borough_data$Category))
ggplot(borough_data, aes(x = Category, y = `Geographic Subdivision`, fill = `% Grads`)) +
  geom_tile(color = "white") +  # Add white grid lines for clarity
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "% Grads") +
  labs(
    title = "Graduation Rates by Borough and Demographic Category",
    x = "Demographic Category",
    y = "Borough"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
)

```

The heatmap provides a borough-wise and demographic-specific breakdown of graduation rates across NYC. Each cell represents a unique combination of borough and demographic category, with darker shades indicating higher graduation rates. This visualization reveals disparities, such as boroughs like Staten Island and Manhattan exhibiting higher graduation rates compared to Bronx. Demographics like "Not ELL" or "Never ELL" often show better outcomes than "Current ELL," indicating potential challenges faced by English learners.

```{r,warning=FALSE}
ggplot(data_clean, aes(x = `# Advanced Regents`, y = `% Grads`)) +
  geom_point(aes(color = `Cohort Year`), size = 1.2, alpha= 0.7, shape = 21, fill = "lightblue", color = "black", stroke = 0.5) +
  theme_minimal() +
  labs(title = "Correlation Between % Advanced Regents and % Grads", 
       x = "% Advanced Regents",
       y = "% Grads")
```
⁠Correlation Between % Advanced Regents and % Grads: This scatter plot illustrates the relationship between the percentage of Advanced Regents diplomas and overall graduation rates. The gradient coloring by cohort year highlights that more recent cohorts tend to have higher concentrations of Advanced Regents diplomas correlating with higher graduation percentages.

```{r,warning=FALSE}
ggplot(data_clean, aes(x = `% Still Enrolled`)) +
  geom_density(fill = "lightblue", alpha = 0.5) +
  geom_density(aes(x = `% Dropout`), fill = "darkblue", alpha = 0.5) +
  theme_minimal() +
  labs(title = "Density Plot of % Still Enrolled vs % Dropout", x = "% Enrolled / Dropout", y = "Density")+
  scale_x_log10()

```
This plot illustrates the distribution of percentages for students still enrolled and those who dropped out. Higher density regions highlight where most students are concentrated in terms of enrollment or dropout rates. For instance, peaks in the density curves indicate common percentage ranges, such as a high number of students either fully enrolled or entirely dropped out. This visualization helps to identify trends in student retention and dropout patterns over time.

```{r}
data_tree <- data_clean |> filter(`Report Category` == "Borough")
treemap(data_tree, 
        index = c("Geographic Subdivision", "Category"), 
        vSize = "# Total Cohort", 
        vColor = "% Grads", 
        draw = TRUE)

```
This treemap shows the distribution of the total cohort across boroughs and demographic categories, such as gender, ethnicity, ELL status, and economic disadvantage. Each rectangle's size reflects the proportion of students in that subgroup, highlighting key differences within and across boroughs. For example, "Not SWD" and "Never ELL" are prominent categories, indicating their larger representation in the cohort.